---
title: "Q6: Class 06 Code Improvement Homework"
author: "STEPHANIE CURTIS"
date: "01/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Removing Calculation Duplication when Coding

Rather than reading into PDB three separate times, making three separate variables, and writing the same three lines of code for each, I have created **a single function** to perform the lookup and analysis, named pdbplotfxn. This function allows the comparison of protein structure under drug-treated and untreated condition (for example) by visualizing the amino acid sequence plotted against the B-factor under both conditions. 


```{r}
#load package
library(bio3d)

#pdbplotfxn outputs a graph of the amino acid sequence along the x-axis to assess what regions of the protein are structurally flexible (B-factor). Different structures from PDB can be input to analyze a protein's structural change in response to treatment, in this case to assess the effect of two drugs on a kinase's structure to see if either can inhibit its ability to bind/phosphorylate substrate, and whether they act on the same region of the protein.
pdbplotfxn <- function(x){
  y <- read.pdb(x)
  sChainA <- trim.pdb(y,chain="A",elety="CA")
                      sb <- sChainA$atom$b
                      chainplot <- plotb3(sb,sse=sChainA,typ="l",ylab="Bfactor")}

#Inputs to the function are the following three PDB ID's in their normal four-letter format as you would input them to a read.pdb function. 
pdbplotfxn("4AKE")
pdbplotfxn("1AKE")
pdbplotfxn("1E4Y")




```


```


